---
title: "Simulation_2023_10_26"
author: "Phuc Quang Vu pqv9968"
date: "2023-10-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading necessary library**

```{r loading library, eval=TRUE,message=FALSE}
library(mvtnorm)
library(multtest)
library(mnonr)

```

**Define function to calculate fwer and fdr given two matrices of hypotheses and testing result**

```{r calculate fwer and fdr, eval=TRUE}
error_cal<-function(null,test){
  #dim(null)=dim(test)=number of repetitions*number of test
  #calculate the total time that null is true and test report a significant result in each replicate
  V<-apply(null==0&test==1,1,sum)
  #calculate by row the total significant test result in each replicate
  R<-apply(test,1,sum)
  #calculate fwer and fdr
  #fwer=P(V>0)
  fwer<-sum(V>0)/length(V)
  #fdr = E(V/R|R>0)*P(R>0)
  fdr<-mean(V[R>0]/R[R>0])*sum(R>0)/length(R)
  return(list('fwer'=fwer,'fdr'=fdr,'v'=V,'r'=R))
}
```



**Define function to generate a random negative orthant dependent positive-definite covariance matrix**
```{r generate random positive-definite covariance, eval=TRUE}
matrix_expander<-function(original){
  d<-dim(original)[1]
  temp<-rep(0,d)
  for (i in 1:d){
    t1<-abs(sum(original[,i])-1)
    t2<-abs(sum(temp))
    t<-max(t1,t2)
    temp[i]<-runif(1,t-1,0)
  }
  new<-cbind(rbind(original,temp),c(temp,1))
  return(new)
}
neg_matrix_gen<-function(begin=-0.1,m=20){
  if (is.matrix(begin)){
    original <-begin
  }
  else {
    original<-matrix(c(1,begin,begin,1),nrow=2)
  }
  
 d<-dim(original)[1]
 for (i in 1:(m-d)){
   original<-matrix_expander(original)
 }
return(original)
}
```




**Define function to return adjusted p values based on mt.rawp2adjp**
```{r}
new_mt.rawp2adjp<-function (rawp, proc = c("Bonferroni", "Holm", "Hochberg", "SidakSS", 
                         "SidakSD", "BH", "BY", "ABH", "TSBH"), alpha = 0.05, na.rm = FALSE) 
{
  m <- length(rawp)
  if (na.rm) {
    mgood <- sum(!is.na(rawp))
  }
  else {
    mgood <- m
  }
  n <- length(proc)
  a <- length(alpha)
  index <- order(rawp)
  h0.ABH <- NULL
  h0.TSBH <- NULL
  spval <- rawp[index]
  adjp <- matrix(0, m, n + 1)
  dimnames(adjp) <- list(NULL, c("rawp", proc))
  adjp[, 1] <- spval
  if (is.element("TSBH", proc)) {
    TS.spot <- which(proc == "TSBH")
    TSBHs <- paste("TSBH", alpha, sep = "_")
    newprocs <- append(proc, TSBHs, after = TS.spot)
    newprocs <- newprocs[newprocs != "TSBH"]
    adjp <- matrix(0, m, n + a)
    dimnames(adjp) <- list(NULL, c("rawp", newprocs))
    adjp[, 1] <- spval
    tmp <- spval
    for (i in (m - 1):1) {
      tmp[i] <- min(tmp[i + 1], min((mgood/i) * spval[i], 
                                    1, na.rm = TRUE), na.rm = TRUE)
      if (is.na(spval[i])) 
        tmp[i] <- NA
    }
    h0.TSBH <- rep(0, length(alpha))
    names(h0.TSBH) <- paste("h0.TSBH", alpha, sep = "_")
    for (i in 1:length(alpha)) {
      h0.TSBH[i] <- mgood - sum(tmp < alpha[i]/(1 + alpha[i]), 
                                na.rm = TRUE)
      adjp[, TS.spot + i] <- tmp*(1+alpha[i]) * h0.TSBH[i]/mgood
    }
  }
  if (is.element("Bonferroni", proc)) {
    tmp <- mgood * spval
    tmp[tmp > 1] <- 1
    adjp[, "Bonferroni"] <- tmp
  }
  if (is.element("Holm", proc)) {
    tmp <- spval
    tmp[1] <- min(mgood * spval[1], 1)
    for (i in 2:m) tmp[i] <- max(tmp[i - 1], min((mgood - 
                                                    i + 1) * spval[i], 1))
    adjp[, "Holm"] <- tmp
  }
  if (is.element("Hochberg", proc)) {
    tmp <- spval
    for (i in (m - 1):1) {
      tmp[i] <- min(tmp[i + 1], min((mgood - i + 1) * spval[i], 
                                    1, na.rm = TRUE), na.rm = TRUE)
      if (is.na(spval[i])) 
        tmp[i] <- NA
    }
    adjp[, "Hochberg"] <- tmp
  }
  if (is.element("SidakSS", proc)) 
    adjp[, "SidakSS"] <- 1 - (1 - spval)^mgood
  if (is.element("SidakSD", proc)) {
    tmp <- spval
    tmp[1] <- 1 - (1 - spval[1])^mgood
    for (i in 2:m) tmp[i] <- max(tmp[i - 1], 1 - (1 - spval[i])^(mgood - 
                                                                   i + 1))
    adjp[, "SidakSD"] <- tmp
  }
  if (is.element("BH", proc)) {
    tmp <- spval
    for (i in (m - 1):1) {
      tmp[i] <- min(tmp[i + 1], min((mgood/i) * spval[i], 
                                    1, na.rm = TRUE), na.rm = TRUE)
      if (is.na(spval[i])) 
        tmp[i] <- NA
    }
    adjp[, "BH"] <- tmp
  }
  if (is.element("BY", proc)) {
    tmp <- spval
    a <- sum(1/(1:mgood))
    tmp[m] <- min(a * spval[m], 1)
    for (i in (m - 1):1) {
      tmp[i] <- min(tmp[i + 1], min((mgood * a/i) * spval[i], 
                                    1, na.rm = TRUE), na.rm = TRUE)
      if (is.na(spval[i])) 
        tmp[i] <- NA
    }
    adjp[, "BY"] <- tmp
  }
  if (is.element("ABH", proc)) {
    tmp <- spval
    h0.m <- rep(0, mgood)
    for (k in 1:mgood) {
      h0.m[k] <- (mgood + 1 - k)/(1 - spval[k])
    }
    grab <- min(which(diff(h0.m, na.rm = TRUE) > 0), na.rm = TRUE)
    h0.ABH <- ceiling(min(h0.m[grab], mgood))
    for (i in (m - 1):1) {
      tmp[i] <- min(tmp[i + 1], min((mgood/i) * spval[i], 
                                    1, na.rm = TRUE), na.rm = TRUE)
      if (is.na(spval[i])) 
        tmp[i] <- NA
    }
    adjp[, "ABH"] <- tmp * h0.ABH/mgood
  }
  list(adjp = adjp, index = index, h0.ABH = h0.ABH[1], h0.TSBH = h0.TSBH[1:length(alpha)])
}
```

**Define function to run simulation with 7 arguments: number of replicate, number of test,correlation and random seed**
```{r, eval=TRUE,message=FALSE}
neg_simutation<-function(replicate=10000,
                         number_of_test=20,
                         begin=-0.1,
                         null_prob=1
                         ){
   # set random seed
  # create an empty result matrix and empty array for each case using fwer or fdr
  r_fwer=matrix(0,nrow=replicate, ncol=number_of_test)  
  nsig_fwer=rep(0,replicate)
  r_fdr=matrix(0,nrow=replicate, ncol=number_of_test) 
  nsig_fdr=rep(0,replicate)
  r_all<- matrix(0,nrow=replicate, ncol=number_of_test)
  nsig_all<-rep(0,replicate)
  #create list of null hypotheses
  null_list<-c(rep(0,ceiling(number_of_test*null_prob)),rep(1,number_of_test-ceiling(number_of_test*null_prob)))
  null_list<-sort(null_list,decreasing = TRUE)
  #create an empty matrix consist of result of null hypotheses
  r_null=matrix(0,nrow=replicate, ncol=number_of_test)
  #generate an array of mean for generate test statistics
  mean_list<-null_list
  # at p(Z>0.841621) of a test statistics ~N(0,1) =0.8, to ensure the new test have a power of 80%
  #the mean to generate test statistics would be 1.96+0.841621
  mean_list[mean_list==1]<-1.96+0.841621
  #create covariance matrix
  cor_ma<-neg_matrix_gen(begin=begin,m=number_of_test)
  
  pb <- txtProgressBar(min = 0, max = replicate, style = 3)
  for (m in 1:replicate){
    # generate test statistics from a multivariate normal distribution 
    #with mean according to mean_list
    
    test_stats<-rmvnorm(n=1,
                        mean=mean_list,
                        sigma = cor_ma)
    
    
    # calculate p values from the test statistics
    p_vals<-2*(1-pnorm(abs(test_stats)))
    #adding null list to matrix of null list by order of p_vals
    r_null[m,]<-null_list[order(p_vals)]
    # calculate adjusted p values 
    m_1<-new_mt.rawp2adjp(p_vals, proc=c("Bonferroni", "Holm", "Hochberg", "SidakSS", "SidakSD")
                      ,alpha = 0.05, na.rm = FALSE)$adjp
    m_2<-p.adjust(p_vals[order(p_vals)],method='hommel')
    m_3<-new_mt.rawp2adjp(p_vals, proc=c("BH", "BY","TSBH"), alpha = 0.05, na.rm = FALSE)$adjp  
    # making a matrix from the adjusted p values 
    r_test_fwer=cbind(m_1[,2:6],m_2)
    r_test_fdr=m_3[,2:4]
    r_test_all=cbind(r_test_fwer,r_test_fdr)
    # making a vector of the number of rejected hypotheses using procedures
    n_test_fwer=apply(r_test_fwer<0.05,2,sum)
    n_test_fdr=apply(r_test_fdr<0.05,2,sum)
    n_test_all=apply(r_test_all<0.05,2,sum)
    #picking the procedure with the most rejected hypotheses
    n_max_fwer=max(n_test_fwer)
    n_max_fdr=max(n_test_fdr)
    n_max_all=max(n_test_all)
    r_fwer[m,]<-r_test_fwer[,which.max(n_test_fwer)]
    r_fdr[m,]<-r_test_fdr[,which.max(n_test_fdr)]
    r_all[m,]<-r_test_all[,which.max(n_test_all)]
    setTxtProgressBar(pb, m)
  }
  # calculate fwer and fdr using r_null and r_fwer, r_fdr 
  result_1<-error_cal(r_null,r_fwer<0.05)
  fwer_fwermethod<-result_1$fwer
  fdr_fwermethod<-result_1$fdr
  result_2<-error_cal(r_null,r_fdr<0.05)
  fwer_fdrmethod<-result_2$fwer
  fdr_fdrmethod<-result_2$fdr
  result_3<-error_cal(r_null,r_all<0.05)
  fwer_all<-result_3$fwer
  fdr_all<-result_3$fdr
  #return result
  
  return(list('fwer_fwermethod'=fwer_fwermethod,'fdr_fwermethod'=fdr_fwermethod,
              'fwer_fdrmethod'=fwer_fdrmethod,'fdr_fdrmethod'=fdr_fdrmethod,
              'fwer_all'=fwer_all,'fdr_all'=fdr_all))
}

```


**Results for random matrix**

```{r, eval=TRUE}
#random matrix = True
print_result_random_no_sep<-function(num_tests=c(5,10,20,50),
                                     begins=c(-0.9,-0.7,-0.5,-0.3,-0.1,-0.01),
                                     null_prob=c(0.1,0.5,1)
                                     )
{
  #create an empty dataframe with 3 columns
  results<-data.frame(matrix(ncol=9,nrow=length(num_tests)*length(null_prob)*length(begins)))
  colnames(results)<-c('#test','null_prob','first correlation',
                       'FWER_fwer_method','FDR_fwer_method',
                       'FWER_fdr_method','FDR_fdr_method',
                       'FWER_all','FDR_all')
  max=length(num_tests)*length(null_prob)*length(begins)
  # create a loop in both number of tests and null_prob

  for (n in 1:length(num_tests)){
    for(p in 1:length(null_prob)){
      for(c in 1:length(begins)){
        #calculate the index of result
        index=(n-1)*length(null_prob)*length(begins)+(p-1)*length(begins)+c
        #assign result
        print(sprintf("progress: %d over %d",index,max ))
        
        result<-neg_simutation(number_of_test = num_tests[n],null_prob=null_prob[p],begin=begins[c])
        results[index,'FWER_fwer_method']<-round(result$fwer_fwermethod,4)
        results[index,'FDR_fwer_method']<-round(result$fdr_fwermethod,4)
        results[index,'FWER_fdr_method']<-round(result$fwer_fdrmethod,4)
        results[index,'FDR_fdr_method']<-round(result$fdr_fdrmethod,4)
        results[index,'FWER_all']<-round(result$fwer_all,4)
        results[index,'FDR_all']<-round(result$fdr_all,4)
        results[index,'#test']<-num_tests[n]
        results[index,'null_prob']<-null_prob[p]
        results[index,'first correlation']<-begins[c]
          
        }
        
      }
      
    }
  
  return(results)
}
```

**Saved result to folder**
```{r}
set.seed(123)
dir="/Users/quangphucvu/Library/CloudStorage/OneDrive-nyu.edu/multiplicity/Results/"
file="negative-orthant-matrix-23-11-20.csv"
result<-print_result_random_no_sep()
write.csv(result,file=paste0(dir,file))
```